Revealing transient strain in geodetic data with Gaussian process regression

### 2017-08-10 RevisedSubmission/hines_hetland_2017_gji.tex
reworded thing

To further explore which covariance function for $T$ best describes SSEs, we compare the observations to the predicted displacements for each covariance function. We consider the data prediction vector to be $\hat{\mitbf{d}} = \left(u(\mitbf{P}) + \mitbf{G}\mitbf{a}\right)|\mitbf{d}_*$. Following a similar procedure as in Section \ref{sec:Method}, it can be shown that $\hat{\mitbf{d}}$ is normally distributed with mean 
\begin{equation}\label{eq:DataPredMean}
\mitbf{\mu}_{\hat{d}} = \left[\begin{array}{cc}
                           C_u(\mitbf{P},\mitbf{P}) & \mitbf{G} \\
                           \end{array}\right]
                     \left[\begin{array}{cc}
                           \mitbf{\Sigma} & \mitbf{G} \\
                           \mitbf{G}^T  & \mitbf{0} \\
                           \end{array}\right]^{-1}
                     \left[\begin{array}{c}
                           \mitbf{d}_* \\
                           \mitbf{0} \\
                           \end{array}\right]
\end{equation}  
and covariance
\begin{equation}\label{eq:DataPredCov}
\mitbf{C}_{\hat{\mitbf{d}}} = C_u(\mitbf{P},\mitbf{P}) - 
                        \left[\begin{array}{cc}
                              C_u(\mitbf{P},\mitbf{P}) & \mitbf{G} \\
                              \end{array}\right]
                        \left[\begin{array}{cc}
                              \mitbf{\Sigma} & \mitbf{G} \\
                              \mitbf{G}^T  & \mitbf{0} \\
                              \end{array}\right]^{-1}
                        \left[\begin{array}{c}
                              C_u(\mitbf{P},\mitbf{P}) \\
                              \mitbf{G}^T \\
                              \end{array}\right].
\end{equation}
We compute $\hat{\mitbf{d}}$ using SE, Wendland, and IBM covariance functions for $T$ and the median hyperparameters from Table 1. Figure \ref{fig:Fit} compares the easting component of $\mitbf{d}_*$ to $\hat{\mitbf{d}}$ during the winter 2015-2016 SSE at the three stations that record the strongest SSE signal, ALBH, P436, and SC02. Based on Figure \ref{fig:Fit}, $\hat{\mitbf{d}}$ is insensitive to the choice of covariance function for $T$. The predicted displacements for the IBM covariance function contain slightly more high frequency, and perhaps spurious, features. Regardless of the choice of covariance function, $\hat{\mitbf{d}}$ at stations ALBH and SC02 appears slightly too smooth when compared to $\mitbf{d}_*$. This could indicate that the chosen timescale hyperparameter, $\tau$, is too large. The predicted displacements at the remaining stations, which record more subtle SSE deformation, seem to faithfully describe $\mitbf{d}_*$.

>>>>

To further explore which covariance function for $T$ best describes SSEs, we compare the observations to the predicted displacements for each covariance function. We consider the data prediction vector to be $\hat{\mitbf{d}} = \left(u(\mitbf{P}) + \mitbf{G}\mitbf{a}\right)|\mitbf{d}_*$. Following a similar procedure as in Section \ref{sec:Method}, it can be shown that $\hat{\mitbf{d}}$ is normally distributed with mean 
\begin{equation}\label{eq:DataPredMean}
\mitbf{\mu}_{\hat{d}} = \left[\begin{array}{cc}
                           C_u(\mitbf{P},\mitbf{P}) & \mitbf{G} \\
                           \end{array}\right]
                     \left[\begin{array}{cc}
                           \mitbf{\Sigma} & \mitbf{G} \\
                           \mitbf{G}^T  & \mitbf{0} \\
                           \end{array}\right]^{-1}
                     \left[\begin{array}{c}
                           \mitbf{d}_* \\
                           \mitbf{0} \\
                           \end{array}\right]
\end{equation}  
and covariance
\begin{equation}\label{eq:DataPredCov}
\mitbf{C}_{\hat{\mitbf{d}}} = C_u(\mitbf{P},\mitbf{P}) - 
                        \left[\begin{array}{cc}
                              C_u(\mitbf{P},\mitbf{P}) & \mitbf{G} \\
                              \end{array}\right]
                        \left[\begin{array}{cc}
                              \mitbf{\Sigma} & \mitbf{G} \\
                              \mitbf{G}^T  & \mitbf{0} \\
                              \end{array}\right]^{-1}
                        \left[\begin{array}{c}
                              C_u(\mitbf{P},\mitbf{P}) \\
                              \mitbf{G}^T \\
                              \end{array}\right].
\end{equation}
We compute $\hat{\mitbf{d}}$ using SE, Wendland, and IBM covariance functions for $T$ and the median hyperparameters from Table 1. Figure \ref{fig:Fit} compares the easting component of $\mitbf{d}_*$ to $\hat{\mitbf{d}}$ during the winter 2015-2016 SSE at the three stations that record the strongest SSE signal, ALBH, P436, and SC02. Based on Figure \ref{fig:Fit}, $\hat{\mitbf{d}}$ is insensitive to the choice of covariance function for $T$. The predicted displacements for the IBM covariance function contain slightly more high frequency, and perhaps spurious, features. When comparing $\hat{\mitbf{d}}$ to $\mitbf{d}_*$ at station ALBH and SC02, it appears that $\hat{\mitbf{d}}$ slightly underestimates the rate of deformation during the SSE, regardless of the chosen covariance function. This over-smoothing could indicate that the chosen time-scale hyperparameter, $\tau$, is too large. The predicted displacements at the remaining stations, which record more subtle SSE deformation, seem to faithfully describe $\mitbf{d}_*$.
